Enhanced Particle Swarm Optimization In Sketch Recognition

### 2010-08-18 Swarm/tmp/swarmModified.tex
removing the comments

 % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
%%% old abstract.. 
 %Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture. This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm. Simulation results will be carried out on stock market indices as an example of time series prediction which shows that our proposed training algorithm is more efficient and accurate than the Basic-PSO in stock market prediction.
\end{abstract}
\pagenumbering{arabic}
 \section{Introduction}

%However, in PSO algorithm as the number of iterations increases, the particles speed decrease leading to converge to the global best point found so far, which is not guaranteed even to be a local minimum.. To enhance the performance and explore better solutions, many researchers introduced GAs operations like crossover, mutation, and selection operators or tuning the PSO parameters leading to increase the computational effort  \cite{Ouyang2009IJCAI}.

%As the need for higher effective neural network architecture that increases the neural network output accuracy is highly required. So, in this study we propose a new architecture based on introducing the hidden layer sigmoid weight which adapts the shape of the hidden layer output sigmoid function. Also, another Enhanced Particle Swarm Optimization training algorithm is applied in training network weights.

@@

%The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive. This is an optimization problem, which evolutionary programming can solve efficiently. A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.

@@

where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

@@

%Otherwise add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm. The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.

%This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
 % \begin{equation}
 % v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
 % \label{eq:ModSwarm}
 % \end{equation}

%where $iRandom_i$ and $Random_j$ are two particles in the solution space. This modification replace the distance between the current particle and the local best with two random particle from the space. The effect of this change in more expolration and avoiding of local min problems. Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated. The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used. 
\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}. The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step. The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency. The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms. Other modifications are explained in the details in the next sections.  
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points. % (as shown are redundant)
\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves. After two different segmentations are generated, the system chooses the segmentation with the minimum error. The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke. % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves.

%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.  
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

\subsection{Non ellipse fitting algorithm}
\label{subsubsec:Discreteparticleswarmalgorithm}
\begin{figure}
	\centering
		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{ppd.jpg}}
	%	\hfill
	 \includegraphics[scale=0.7]{pso1.jpg}			
	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}
Two DPSO algorithms are used to generate different segmentations for each stroke the user draws. For each segmentation generated, an error is evaluated. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. 
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

@@

As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

@@

		\includegraphics[scale=0.5]{symbolSet.PNG} 
		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
	 \end{figure}
	  %The first experiment with the whole feature set and the second experiments with only half of the features.
To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set. The second experiment was tested on both algorithm but only using half the feature set. Both experiments were compared with the orginal sketch recognition system prsenented in \cite{mypaper}.

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the sytem using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.   %n the We performed two experiments to test the system. Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization. To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.

%In our experiment we used three different datasets. First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB). The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. We collected the second and third datasets. Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.  Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments. The datasets were divided into training set and test set. Five different splits  for the training and test data are generated from each dataset. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.

 %The goal of the system is to modify the segmentation and recognition of the sketch system. It made the same type of experiement done in \cite{mypaper}. The dataset is divided into 13 category of goemetric %symbols

% The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.  Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The datasets were divided into training and test sets. Five different splits for the training and test data are generated from each dataset. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both the orginal system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increases the performance over the orginal system. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both the orginal system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increase the recognition rate. }
   \label{exp1}
\end{figure}

% We performed two experiments to test the system. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.

@@

%The training and testing performance measurements using both PSO and Enhanced-PSO in training and testing phases are shown in figure\ref{exp2} and figure \ref{exp1}.

@@

 For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm. Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.   %this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification

 %A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.

>>>>

\end{abstract}
\pagenumbering{arabic}
 \section{Introduction}

@@

@@

where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.

 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

@@

\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}. The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step. The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency. The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms. Other modifications are explained in the details in the next sections.  
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points. % (as shown are redundant)
\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves. After two different segmentations are generated, the system chooses the segmentation with the minimum error. The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}

This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

\subsection{Non ellipse fitting algorithm}
\label{subsubsec:Discreteparticleswarmalgorithm}
\begin{figure}
	\centering

	 \includegraphics[scale=0.7]{pso1.jpg}			
	\caption{DPSO algorithm encoding}
	\label{fig:pso1}
\end{figure}
Two DPSO algorithms are used to generate different segmentations for each stroke the user draws. For each segmentation generated, an error is evaluated. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. 
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

@@

As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}

\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

@@

		\includegraphics[scale=0.5]{symbolSet.PNG} 
		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
	 \end{figure}

To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set. The second experiment was tested on both algorithm but only using half the feature set. Both experiments were compared with the original sketch recognition system presented in \cite{mypaper}.

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the system using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.   
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both the original system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increases the performance over the original system. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both the original system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increase the recognition rate. }
   \label{exp1}
\end{figure}

@@

@@

 For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm. Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.
